---
title: "Reproducible Research: Peer Assessment 1"

---
***
PLEASE NOTE:: In order to View the Graphs associated with this assignment, you must look at the HTML posting.  Thanks.. 
***

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


## Loading and preprocessing the data
```{r}
library(dplyr)
setwd("C:/Users/uszlj8j/Desktop/Learning R/Data Science CourseRA/R_Reproducable_Research")
Activity <- read.csv("activity.csv", header = TRUE)

View(head(Activity))
Activity$steps <- as.numeric(Activity$steps)
Activity$interval <- as.numeric(Activity$interval)

```




## What is mean total number of steps taken per day?
```{r, echo=TRUE}

#Calculate the number of Steps per day

Nbr_steps <- select(Activity, date, steps) %>% group_by(date) %>% summarise(Steps_per_day = sum(steps))

hist(Nbr_steps$Steps_per_day)

#Mean and Standard Deviation of steps per day

Mean_Median_steps <- Nbr_steps %>% summarise(Mean_Steps_per_day = mean(Steps_per_day, na.rm = TRUE), 
                                           Median_Steps_per_day = median(Steps_per_day, na.rm = TRUE))

Mean_Median_steps
```



## What is the average daily activity pattern?
```{r, echo=TRUE}

data <- select(Activity,interval, steps) %>% group_by(interval) %>% 
                        summarise(Avg_Steps_per_interval = mean(steps, na.rm = TRUE))

plot( x=data$interval, y=data$Avg_Steps_per_interval, type="l")


#Which interval contains max number of steps
# Answer: Interval 835 by looking at the output dataset

```



## Imputing missing values
```{r}

#Count the number of NAs
Steps_NA <- sum(is.na(Activity$steps))
Steps_NA


#Impute Missing Values, use Mean for the Interval to impute missing values
#Create a new dataframe
Activity_Impute <- Activity %>% group_by(interval) %>% mutate(ifelse(is.na(steps),mean(steps, na.rm = TRUE),steps))
colnames(Activity_Impute)[4] = "Steps_Imputed"


#Calculate the number of Steps per day

Nbr_steps_Impute <- select(Activity_Impute, date, Steps_Imputed) %>% group_by(date) %>% summarise(Steps_per_day = sum(Steps_Imputed))

hist(Nbr_steps_Impute$Steps_per_day)


#Calculate Mean and Median Steps per day

Mean_Median_steps_Impute <- ungroup(Activity_Impute) %>% select(Steps_Imputed)  %>% 
                            summarise(Mean_Steps_per_day = mean(Steps_Imputed), 
                                    Median_Steps_per_day = median(Steps_Imputed))

Mean_Median_steps_Impute

```



## Are there differences in activity patterns between weekdays and weekends?
```{r}

#Convert date variable into Date funciton, create a new variable that gives the weekday

Activity_Impute$date <- as.Date(Activity_Impute$date)

Activity_Impute <- Activity_Impute %>% mutate(weekdays = weekdays(date))
Activity_Impute <- Activity_Impute %>% mutate(Week_Type = ifelse((weekdays %in% c("Saturday", "Sunday")),"Week_End", "Week_Day"))

Week_day <- select(Activity_Impute,interval, Steps_Imputed, Week_Type) %>% filter(Week_Type == "Week_Day") %>%
             group_by(interval, Week_Type) %>% summarise(Avg_Steps_per_interval = mean(Steps_Imputed, na.rm = TRUE))

Week_End <- select(Activity_Impute,interval, Steps_Imputed, Week_Type) %>% filter(Week_Type == "Week_End") %>%
             group_by(interval, Week_Type) %>% summarise(Avg_Steps_per_interval = mean(Steps_Imputed, na.rm = TRUE))

 
par(mfrow=c(2,1)) 
plot(x= Week_day$interval, y=Week_day$Avg_Steps_per_interval, type = "l")
plot(x= Week_End$interval, y=Week_End$Avg_Steps_per_interval, type = "l")

```

